---
title: "potatoheads"
author: "Ann E. Nordmeyer"
date: "November 17, 2015"
output: html_document
---

Analysis code for potato-or study (Ann E. Nordmeyer & Michael C. Frank).  Extension of book-or, but with potato heads. Experiment can be viewed at http://langcog.stanford.edu/expts/AEN/potatoes/potatoes.html

Participants saw two sets of potato heads (each set contained 2 potato heads) and read a sentence ("Find the potatoes with...").  Potatoes had either one or two features (there were six total features: teeth, arms, ears, hair, glasses pants).  Participants were asked to figure out which potatoes the sentence described.

Three sentence types (word.type in data frame): 

* and ("Find the potatoes with X and Y")

* or ("Find the potatoes with X or Y")

* noun ("Find the potatoes with X")


Three types of "book sets": 

* and (both potatoes had both X and Yl)

* or (one potato has just X, one potato has just Y)

* noun (both potatoes have just X)


Combining these types of book sets gives us three possible combinations (image.type in data frame): 

* and/or

* and/noun

* or/noun


The experiment is fully crossed, leaving us with 9 trial types.  Participants saw four trials per trial type for a total of 36 trials.  

# Setting up

Load required Libraries
```{r libraries}
library(dplyr)
library(ggplot2)
library(tidyr)
library(bootstrap)
library(knitr)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

Some useful functions:
```{r functions}
## number of unique subs
n.unique <- function (x) {
  length(unique(x))
}

## for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
ci.high <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.975)}
```

Load in data, get demographic info, and clean up

```{r data}
#Load in data
d <- read.csv("../long_data/potatoes_long_mturk.csv") 

n.original <- n.unique(d$subid)

#Sometimes turkers say they are under 18, which they aren't supposed to be if they are on Turk, but I exclude them anyway
d <- filter(d, age != 17)
n.age.exclusions <- n.original - n.unique(d$subid)
  
#what languages?
d <- filter(d, language == "English")
n.lang.exclusions <- n.original - n.age.exclusions- n.unique(d$subid)

n.final <- n.unique(d$subid)

d <- select(d, subid, trial.num, item, word.type, image.type, response, rt)
d$word.type <- factor(d$word.type, levels = c("and","or","noun"))
d$image.type <- factor(d$image.type, levels = c("and/or", "and/noun", "or/noun"))
```

We originally collected `r n.original` participants.  `r n.age.exclusions` were excluded for indicating that they are under 18 years of age (our introduction slide clarifies that people are not eligible to take the study if they are under 18, and you aren't supposed to have an MTurk account if you are under 18, but a few people still make this selection anyway).  `r n.lang.exclusions` were excluded for indicating that English is not their primary language.  This left a total of `r n.final` participants.  


# Response analysis

What do participants select for each trial type?

```{r choice}

ms.response <- d %>%
  group_by(subid, word.type, image.type, response) %>%
  summarize(n = n()) %>%
  spread(response, n) %>%
  replace(is.na(.), 0) %>% 
  mutate(total = and + noun + or) %>%
  gather(response, n, and:or) %>%
  mutate(prop = n / total) %>%
  group_by(word.type, image.type, response) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop)) %>%
  mutate(predictions = ifelse(word.type == "and", "semantics", "unclear"), 
         predictions = ifelse((word.type == "or" & image.type == "and/or") | (word.type == "or" & image.type == "or/noun"), "scalar", predictions), 
         predictions = ifelse((word.type == "noun" & image.type == "and/noun") | (word.type == "noun" & image.type == "or/noun"), "ad-hoc", predictions)) %>%
  filter((image.type == "and/noun" & response != "or") |
           (image.type == "and/or" & response != "noun") |
           (image.type == "or/noun" & response != "and"))

qplot(data = ms.response,
      x = response, y = prop, fill = predictions,
      geom = "bar", stat = "identity", position = "dodge") + 
      facet_grid(word.type ~ image.type, scales = "free") + 
  geom_errorbar(aes(ymin = cil, ymax = cih), 
                position = position_dodge(.9), width = 0, ) +
  theme_bw()
```

# Reaction time

How long does it take participants on each trial type?

I'm not sure this is actually an interesting analysis, but I had the RT data so here it is.  It's messy because participants read sentences, and presumably there's a lot of variability in reading time.

```{r rt}
ms.rt <- d %>%
  filter(log(rt) < mean(log(rt) + 3 * sd(log(rt))), 
         log(rt) > mean(log(rt) - 3 * sd(log(rt)))) %>%
  group_by(subid, word.type, image.type) %>%
  summarize(rt = mean(rt)) %>%
  group_by(word.type, image.type) %>%
  summarize(cih = ci.high(rt),
            cil = ci.low(rt),
            rt = mean(rt))


qplot(data = ms.rt, x = image.type, y = rt, fill = word.type, 
      geom = "bar", stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = cil, ymax = cih), 
                position = position_dodge(.9), width = 0) +
  theme_bw()
```
